### 2018-10-26 reports/manuscript/sections.tex
Update permissions

\input{sections/introduction}
\input{sections/methods}
\input{sections/results}
\input{sections/conclusion}

\section{Acknowledgements}

Here are some amazing acknowledgements.