### 2019-02-07 reports/manuscript/sections.tex
Add methods section to manuscript

\input{sections/introduction}
\input{sections/methods}
\input{sections/results}
\input{sections/conclusion}

\section*{Acknowledgments}

Here are some amazing acknowledgements.