### 2019-11-25 reports/manuscript/sections.tex

\input{sections/introduction}
\input{sections/methods}
\input{sections/results}
\input{sections/conclusion}

\section*{Acknowledgments}

This work was supported by BRAIN Iniative grant 1RF1MH121868-01 from the
National Institutes for Mental Health and by a grant from the Gordon \& Betty
Moore Foundation and the Alfred P. Sloan Foundation to the University of
Washington eScience Institute Data Science Environment. We would like to thank
Scott Murray for a useful discussion of the SGL method.