### 2019-12-03 reports/manuscript/sections/methods.tex
Add details to cross-validation section and software implementation. Update some references and minor edits to other sections

\section*{Materials and methods}

\subsection*{Data}

Two different previously-published datasets were used here:

\begin{enumerate}

\item Diffusion MRI data from a previous study of properties of the white matter
across the lifespan\cite{yeatman2014lifespan}, containing dMRI data from 76
subjects with ages 6-50. These data were measured in a GE Discovery 750 3T MRI
scanner at the Stanford Center for Cognitive and Neurobiological Imaging. The
Stanford University IRB approved the procedures of this study. Informed consent
was obtained from each adult participant, and assent for participation was
provided by parents/guardians for children. Voxel resolution was
\num{2x2x2}\si{\milli\meter^3} with 96 non-colinear directions measured with a
$b= \SI{2000}{\second\per\milli\meter^2}$ and 30 non-colinear directions
measured with a $b=\SI{1000}{\second\per\milli\meter^2}$.
Data was preprocessed to correct for subject motion and the diffusion
tensor model \cite{basser1994mr} was fit in every voxel, using a robust fit
\cite{chang2005restore}. These data were acquired using a dual spin echo
sequence, in which there is sufficient time for eddy currents to subside between
the application of the gradients and the image acquisition, so no eddy current
correction was appliced. We will refer to this dataset as WH.

\item Diffusion MRI from from a previous study of the corticospinal tract
profile in patients with amyotrophic lateral sclerosis
(ALS)\cite{sarica2017corticospinal}, containing data from 24 ALS patients and 24
demographically matched healthy controls. These data were measured in a GE
Discovery 750 3T MRI scanner at the Institute of Bioimaging and Molecular
Physiology in Catanzaro. Informed consent was provided as approved by the
Ethical Committee of the University ``Magna Graecia'' of Catanzaro. Voxel
resolution was \num{2x2x2}\si{\milli\meter^3} and 27 non-colinear directions
were measured with a $b=\SI{1000}{\second\per\milli\meter^2}$.
Here, both motion correction and eddy current correction were
applied before modeling the diffusioon tensor in each voxel. We will refer to
this dataset as ALS.

\end{enumerate}

Data from both of these studies was processed in a similar manner, using the
Matlab Automated Fiber Quantification toolbox (AFQ) \cite{yeatman2012tract}:
streamlines representing fascicles of white matter nerve fibers were generated
using a determinstic tractography algorithm that follows the prinicpal diffusion
direction of the diffusion tensor in each voxel (STT \cite{basser2000vivo}).
Major tracts were identified using multiple criteria: streamlines are selected
as candidates for inclusion in a bundle of streamlines that represents a tract
if they pass through known inclusion ROIs and do not pass through exclusion ROIs
\cite{wakana2007reproducibility}. In addition, a probabilistic atlas is used to
exclude streamlines which are unlikely to be part of a tract \cite{Hua2008-sh}.
Each streamline is resampled to 100 nodes and the robust mean at each location
is calculated by estimating the 3D covariance of the location of each node and
excluding streamlines that are more than 5 standard deviations from the mean
location in any node. Finally, a tract profile of tissue properties in each
tract was created by interpolating the value of MRI maps of these tissue
properties to the location of the nodes of the resampled streamlines designated
to each tract. In each of 100 nodes, the values are summed across streamlines,
weighting the contribution of each streamline by the inverse of the mahalnobis
distance of the node from the average  of that node across streamlines. This
means that streamlines that are more representative of the tract contribute more
to the tract profile, relative to streamlines that are on the edge of the tract.

This process creates tract profiles, in which diffusion measures are quantified
and averaged along twenty major fiber tracts. Here, we use only the mean
diffusivity (MD) and the fractional anisotropy (FA) of the diffusion tensor, but
additional dMRI-based maps or maps based on other quantitative MRI measurements
can also be used. These tract profiles, along with the phenotypical data we wish
to explain or predict, form the input to the SGL algorithm. In a domain-agnostic
machine learning context, the phenotypical data comprise the target variables
while the tract profiles form the feature or predictor variables (See
Fig~\ref{fig:group-structure}).

\subsection*{Sparse Group Lasso}

Before fitting a model to the data, imputation and standardization are
performed. Missing node values (e.g., in cases where AFQ designates a node as
not-a-number) are imputed via linear interpolation. Care is taken not to
interpolate across the boundaries between different tracts. Some diffusion
metrics will have naturally larger variance than others and may therefore
dominate the objective function and make the SGL estimator unable to learn from
the lower variance metrics. For example, fractional anisotropy (FA) is bounded
between zero and one and could be overwhelmed by an unscaled higher-variance
metric like the mean diffusivity (MD). To prevent this, we use the
\lstinline{StandardScaler} from scikit-learn\cite{scikit-learn}, to remove each
feature's mean and scale it to unit variance. Scaling is performed separately
within each cross-validation set's training or testing data to prevent leakage
of information between the testing and training sets\cite{kaufman2012leakage}.

After scaling and imputation, the tractometry data and target phenotypical data
can be organized in a linear model:

\begin{equation}
y = \mathbf{X} \cdot \beta,
\label{eq:lm}
\end{equation}

where $y$ is the phenotype -- categorical, such as a clinical diagnosis, or
continuous numerical, such as the subject's age. The tractometry data is
represented in the feature matrix $\mathbf{X}$, with rows corresponding to
different subjects, and columns corresponding to diffusion measures at different
nodes within each tract. The relationship between tractometric features and the
phenotypic target is characterized by the coefficients in $\beta$. In general,
the feature matrix $\mathbf{X}$ has dimensions $S \times (T \cdot N \cdot M)$,
where $S$ is the number of subjects, $T$ is the number of white matter tracts,
$N$ is the number of nodes in each tract, and $M$ is the number of diffusion
metrics calculated at each node. Typically, $T = 20$, $N = 100$, and $2 \le M
\le 8$, resulting in $\sim 4,000 - 16,000$  features. Conversely, many dMRI
studies have between a few dozen and a few hundred subjects, yielding a feature
matrix that is wide and short. Even in cases where more than a thousand subjects
are measured (e.g., in the Human Connectome Project, where 1,200 subjects were
measured \cite{VanEssen2012}), the problem is ill-posed: the high dimensionality
of this data requires regularization to avoid overfitting and generate
interpretable results.

Here, we propose that in addition to regularizing the coefficients in $\beta$,
we can also capitalize on our knowledge of the group structure of the tract
profile features in $\mathbf{X}$. The tract-metric combinations form a natural
grouping. For example, we expect that MD features within the left arcuate
fasciculus will co-vary across individuals. Likewise for FA values within the
right corticospinal tract (CST) and so on. This group structure is represented
in Fig~\ref{fig:group-structure}, which depicts the linear model $y = \mathbf{X}
\cdot \beta$. Thus, we seek a regularization approach that will fit a linear
model with anatomically-grouped covariates, where we expect to observe both
groupwise sparsity, where the number of groups (tract/metric combinations) with
at least one non-zero coefficients is small, as well as within-group sparsity,
where the number of non-zero coefficients within each non-zero group is small.
The sparse group lasso (SGL)
is a penalized regression technique that satisfies exactly these
criteria\cite{simon2013sparse}. It solves for a coefficient vector
$\widehat{\beta}$ that satisfies
\begin{equation}
    \widehat{\beta} = \min_\beta \frac{1}{2}
    \norm*{y - \displaystyle \sum_{\ell = 1}^{G}
    \mathbf{X}^{(\ell)} \cdot \beta^{(\ell)}}_2^2
    + \lambda_1 \displaystyle \sum_{\ell = 1}^{G}
    \sqrt{p_\ell} \norm*{\beta^{(\ell)}}_2
    + \lambda_2 \norm*{\beta}_1,
    \label{eq:sgl}
\end{equation}

where $G$ is the number of groups $\mathbf{X}^{(\ell)}$ is the submatrix
of $\mathbf{X}$ corresponding to group $\ell$, $\beta^{(\ell)}$ is
the coefficient vector for group $\ell$ and $p_\ell$ is the length of
$\beta^{(\ell)}$. In the tractomtetry setting, $G = T \cdot M$ and
$\forall \ell: p_\ell = 100$. The first term is the mean square error
loss, $L_{\text{mse}}$, as in the standard linear regression framework.
The second and third terms encourage groupwise sparsity and overall
sparsity, respectively. If $\lambda_1 = 0$ and $\lambda_2 = 1$, the
SGL reduces to the traditional lasso\cite{tibshirani1996regression}.
Conversely, if $\lambda_1 = 1$ and $\lambda_2 = 0$, the SGL reduces to
the group lasso\cite{yuan2006model}.

\subsection*{Implementation, cross-validation and metaparameter optimization}

For given values of $\lambda_1$ and $\lambda_2$, the cost function in equation
\ref{eq:sgl} can be optimized using proximal gradient descent methods
\cite{parikh2014proximal} here implemented as a custom proximal operator that is
then optimized using the C-OPT library\cite{copt}. This supplies an estimate
of the optimal $\widehat{\beta}$ given a particular set of values for the
regulation parameters $\lambda_1$ and $\lambda_2$. We then find the optimal
values for $\lambda_1$ and $\lambda_2$ using the hyperopt library's \verb|fmin|
function\cite{Bergstra_2015} with a tree of Parzen estimators search
algorithm\cite{bergstra2011algorithms}. For the WH dataset, where the target
variable is continuous numerical, \verb|fmin| searches for metaparameters that
minimize the median absolute error, though in general a user may also
choose to minimize the root-mean-square-error or to maximize the
coefficient of determination ($R^2$). For the ALS dataset, where the
target variable is binary categorical, \verb|fmin| seeks to maximize the
classification accuracy, though in general the user may also choose to
maximize the area under the receiver operating curve (ROC AUC) or the
average precision. \adam{Enough detail provided?}

To objectively evaluate the model and guard against over-fitting,
we used a nested cross-validation scheme, depicted in
Fig~\ref{fig:cross-val} for the categorical classification case. The
input data is shuffled and then decomposed into $k$ batches, hereafter
referred to as folds. For the ALS dataset, $k=10$, while for the WH
dataset, $k=5$. For each unique fold, we hold that fold out as the
test\textsubscript{outer} set and let the remaining data comprise
the train\textsubscript{outer} set, with the subscript indicating
the depth of the nested decomposition. We further decompose each
train\textsubscript{outer} set into three folds, and again for each
unique fold, we hold out that fold as the test\textsubscript{inner}
set and let the remaining data comprise the train\textsubscript{inner}
set. At level-1 of the decomposition, we fit an SGL model using
arbitrary fixed regularization parameters $\lambda_1$ and $\lambda_2$,
training the model using train\textsubscript{inner} and evaluating
the fit using the appropriate metric on test\textsubscript{inner}.
The optimal regularization parameters are determined using the
aforementioned hyperopt library at level-0 of the decomposition. We
find optimal regularization parameters and $\widehat{\beta}$ for each
train\textsubscript{outer} set and then use those to predict values for
data in test\textsubscript{outer}. Thus each subject in the dataset has
a predicted phenotype derived from a model that never saw its particular
subject's data.

The above procedure describes $k$-fold cross validation. In fact,
we use repeated $k$-fold cross validation on the outer level of the
decomposition, so that the input data is decomposed into $k$ folds,
three times. Thus, each subject has three predicted phenotypes. We then
take the mean predicted value to summarize the skill of the model.
In the classification case, the shuffling into folds is stratified such
that each fold has a population that preserves the percentage of each
class found in the larger input data.
\adam{Clear as mud?}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.55\textwidth]{nested-cross-validation.pdf}
    \caption{{\bf Nested $k$-fold cross-validation scheme.}
        We evaluate model quality using a nested $k$-fold cross
        validation scheme. At level-0, the input data is decomposed
        three times into $k$ shuffled groups and optimal hyperparameters
        are found for the level-0 training set. Optimization of these
        hyperparameters requires the use of the hyperopt library and
        many repeated evaluations of an SGL model over a search space
        of possible regularization parameters. These evaluations take
        place at level-1 of the decomposition, where the level-0 training
        set is further decomposed into three shuffled groups.
        For the ALS data, $k=10$. For the WH data, $k=5$.
    }
    \label{fig:cross-val}
\end{figure}

\subsubsection*{Incorporating transformations on $y$}

Often, the target variable $y$ is not in the domain in which the linear model
can be best fit to it. Equation \ref{eq:lm} can be slightly modified as:
\begin{equation}
y = f^-1(\mathbf{X} \cdot \beta),
\label{eq:lm}
\end{equation}
where the transformation function $f^{-1}$ characterizes the transform applied
to the data before fitting the linear coefficients. For example, an often-used
transform is the logarithmic transform:
\begin{equation}
f(y) = \log_n(y)
\label{eq:log_nonlinearity}
\end{equation}
In this case, the model is parametrized by one additional fit parameter, $n$.

\subsubsection*{Classification of categorical $y$}
When the phenotypical target variable is categorical, as in the case of
explaining or predicting the presence of a clinical diagnosis, the SGL is
readily adapted to logistic regression, where the probability of a target
variable belonging to an arbitrary defined ``true'' class is the logistic
function of the result of the linear model,
\begin{equation}
    p(y = 1) = \frac{1}{1 + \exp(\mathbf{X}\cdot \beta)},
    \label{eq:logit}
\end{equation}
or equivalently, the mean squared error loss function in Eq~\eqref{eq:sgl} is
replaced with the cross-entropy loss, which for binary classification is the
negative log likelihood of the SGL classifier giving the ``true'' label:
\begin{equation}
    L_{\text{mse}} \rightarrow L_{\log} =
    -\left(y \log(p) + (1 - y) \log(1 - p)\right).
    \label{eq:logloss}
\end{equation}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.65\textwidth]{dMRI_group_structure.png}
    \caption{{\bf dMRI group structure.}
        The phenotypical target data and tractometric features can be organized
    into a linear model, $y = \mathbf{X} \cdot \beta$. The feature matrix
    $\mathbf{X}$ is color-coded to reveal a natural group structure: the left
    (orange) group contains $k$ features from the inferior fronto-occipital
    fascicle (IFOF), the middle (green) group contains $k$ features from the
    corpus callosum, and the right (blue) group contains $k$ features from the
    uncinate. The coefficients in $\beta$ follow the same natural grouping.
    Fascicle image reproduced with permission from Ref~\cite{yeatman2012tract}
    Figure 1.}
    \label{fig:group-structure}
\end{figure}

\subsection*{Software implementation}

The full software implementation of the SGL approach presented here is available
in a Python software package called AFQ-Insight, which is developed publicly in
\url{https://github.com/richford/afq-insight}. The version of the code used to
produce the results herein is also available in \ariel{Need to add Zenodo DOI}.
AFQ-Insight reads the target and feature data that has been processed by AFQ
from comma separated value (CSV) files conforming to the AFQ-Browser data
format\cite{yeatman2018browser} and represents them internally as
\lstinline{DataFrame} objects from the pandas Python
library\cite{mckinney2010data}. The software provides different options for
imputing missing data in the feature matrix. Missing interior nodes are imputed
using linear interpolation. For missing exterior nodes, the user may choose
between linear extrapolation and constant forward(back)-fill. Imputation uses
only values from adjacent nodes in the same white matter tract in the same
subject so there is no danger of data leakage from other subjects. It uses the
scikit-learn\cite{scikit-learn} library to decompose input data into separate
test and train datasets, to scale each feature to have zero mean and
unit variance, and to evaluate each fit in the hyperparameter search using
appropriate classification and regression metrics such as accuracy, area
under the receiver operating curve (AUC ROC), and coefficient of determination
($R^2$). For each set of hyperparameters, we solve the SGL using a custom
proximal operator supplied to the C-OPT library\cite{copt}. Appropriate
hyperparameters are found using the hyperopt library\cite{Bergstra_2015}.