### 2019-12-12 reports/manuscript/sections/conclusion.tex
Incorporate Jason's round 2 edits

\section*{Conclusion}

We present here a novel method for analysis of dMRI tractometry data that relies
on the Sparse Group Lasso \cite{simon2013sparse} to (a) make accurate
predictions of phenotypic properties of individual subjects while,
simultaneously, (b) identifying the features of the white matter that are most
important for this prediction in a completely data-driven approach. The method
is broadly applicable to a wide range of research questions: it performs well in
predicting both continuous variables, such as biological age, as well as
categorical variables, such as whether a person is a patient or a healthy
control. In both of these cases, SGL out-performs previous algorithms that have
been developed for these tasks \cite{sarica2017corticospinal, Richard2018-ux}.
The nested cross-validation approach used to fit the model and make both
predictions and inferences from the model guards against overfitting and tunes
the degree of sparseness required by the algorithm. This means that SGL can
accurately describe phenomena that are locally confined to a particular
anatomical location or diffusion property (e.g., FA in the CST) as well as
phenomena that are widely distributed amongst brain regions and measured
diffusion properties.

Specifically, we demonstrated that the algorithm correctly detects
the fact that ALS, which is a disease of lower motor neurons, is
localized to the cortico-spinal tract. This recapitulates the results
of previous analysis of these same data, using a targeted ROI-based
approach \cite{sarica2017corticospinal}. In contrast, for the analysis
of biological age, the coefficients identified by the algorithm are very
widely distributed across many parts of the white matter, mirroring
previous results with this dataset (and others) that show a large and
continuous distribution of life-span changes in white matter properties
\cite{yeatman2014lifespan}.

Taken together, these results demonstrate the promise of the group-regularized
regression approach. Even at the scale of dozens of subjects, the results
provided by SGL are both accurate, as well as interpretable
\cite{Murdoch2019-ax}: tractometry capitalizes on domain knowledge to engineer
meaningful features; SGL scores these features based on their relative
importance; enables a visualization of these feature importance scores in the
anatomical coordinate frame of the tracts (e.g., Figures \ref{fig:class-results}
and fig:regress-beta) and provides a means to understand model errors (e.g.,
Figure \ref{fig:class-errors}) . Thus, this multivariate analysis approach both
(a) achieves high cross-validated accuracy for precision medicine applications
of dMRI data and (b) identifies relevant features of brain anatomy that can
further our neuroscientific understanding of clinical disorders.

Neuroscience has entered an era in which consortium efforts are putting together
large datasets of high-quality dMRI measurements to address a variety of
scientific questions \cite{jernigan2016ping, jernigan2018abcd,
alexander2017open, Miller2016-hw, VanEssen2012}. The volume and complexity of
these data pose a substantial challenge. Dimensionality reduction with
tractometry, followed by analysis with the approach we present here promises to
capitalize on the wealth of data and on the co-measurement of interesting and
important phenotypical data about brain health and about the participants'
cognitive abilities. We also expect the group-regularized approach to improve
with larger datasets.

SGL has many other potential applications in neuroscience, because
of the hierarchical and grouped nature of many data types that are
collected in multiple sample points within anatomically-defined areas.
For example, this method may be useful to understand the relationship
between fMRI recordings and behavior, where activity in each voxel
may co-vary with voxels within the same anatomical region and form
features and groups of features. Similarly, large-scale multi-electrode
recordings of neural activity in awake behaving animals are becoming
increasingly feasible \cite{steinmetz2018distributed, Jun2017-gv} and
these recordings can form features (neurons) and groups (neurons within
an anatomical region). More ambitiously perhaps, this approach may be
used to understand the role of correlations in so-called resting-state
fMRI time-series and behavior, where pairwise correlations between
voxels in different anatomical regions are features in the matrix and
features may be grouped by pairs of anatomical regions. Given the large
number of voxels in the surface of the gray matter and given that
correlations increase the number of features by a factor of $n^2$, this
would pose a challenging problem to solve using SGL.

The results we present here also motivate extensions of the method using more
sophisticated cost functions. For example, the fused sparse group lasso (FSGL)
\cite{zhou2012} extends SGL to enforce additional spatial structure: smoothness
in the variation of diffusion metrics along the tracts. As brain measurements
include additional structure (for example, bilateral symmetry), future work
could also incorporate overlapping group membership for each entry in the tract
profiles \cite{Rao2014-xm}. For example, a measurement could come from the
corpus callosum, but also from the right hemipshere. This would also require
extending the cost function used here to incorporate these constraints.
Similarly, unsupervised dimensionality reduction of tractometry data (e.g.,
\cite{Chamberland2019-mu}) could also benefit from constraints based on
grouping.

The method is packaged as open-source software called AFQ-Insight
that is openly available, and provides a clear API to allow
for extensions of the method. The sofware integrates within a
broader automated fiber quantification software ecosystem: AFQ
\cite{yeatman2012tract}, which extracts tractometry data from raw and
processed dMRI datasets, as well as AFQ-Browser, which visualizes
tractometry data and facilitates sharing of the results of dMRI studies
\cite{yeatman2018browser}. To facilitate reproducibility and ease use
of the software, the results presented in this paper are also provided
in \url{https://github.com/richford/afq-insight-paper} as a series of
Jupyter notebooks \cite{kluyver2016jupyter}.