### 2019-12-12 reports/manuscript/abstract.tex
Change order of ALS and aging studies to match the order used in the results section

\section*{Abstract}

The white matter contains long-range connections between different
brain regions and the organization of these connections hold important
implications for brain function in health and disease. Tractometry
uses diffusion-weighted magnetic resonance imaging (dMRI) data
to quantify tissue properties (e.g. fractional anisotropy (FA),
mean diffusivity (MD), etc.), along the trajectories of these
connections\cite{yeatman2012tract}. Statistical inference from
tractometry data usually either (a) averages these quantities along
the length of the tract in each individual, or (b) performs analysis
point-by-point along each tract, with group comparisons or regression
models computed separately for each point along every one of the tracts.
In the present work, we developed a method based on the sparse group
lasso (SGL) \cite{simon2013sparse} that takes into account tissue
properties measured along all of the tracts, and selects informative
features by enforcing sparsity, not only at the level of individual
tracts, but also across the entire set of tracts and all of the measured
tissue properties. The sparsity penalties for each of these constraints
is identified using a nested cross-validation scheme that guards
against over-fitting and simultaneously identifies the correct
level of sparsity. We demonstrate the accuracy of the method in two
settings: in a classification setting, patients with amyotrophic
lateral sclerosis (ALS) are accurately distinguished from matched
controls\cite{sarica2017corticospinal} and SGL automatically identifies
FA in the corticospinal tract as important for this classification --
correctly finding the parts of the white matter known to be affected
by the disease. In a regression setting, dMRI is used to accurately
predict ``brain age'' \cite{yeatman2014lifespan, Brown2012-so}. In this
case, the weights are distributed throughout the white matter indicating
that many different regions of the white matter change with development
and contribute to the prediction of age. Thus, SGL makes it possible
to leverage the multivariate relationship between diffusion properties
measured along multiple tracts to make accurate predictions of subject
characteristics while simultaneously discovering the most relevant
features of the white matter for the characteristic of interest.