### 2021-01-21 reports/manuscript/sections/methods.tex

\subsection*{Methods}

Methods and associated references are available in \cref{supp:sec:methods} of
the supplemental material \cite{supplement}.