### 2021-01-25 reports/manuscript/sections/methods.tex

\subsection*{Methods}

Methods and associated references are available in the supplemental material
\cite{supplement}.